Storm surges in the Strait of Georgia simulated with a regional model

### 2015-01-19 AO_guides/manuscriptAO.tex
one last read through

%overview oF SoG/Salish Sea models
The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion. %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers. Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome. They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004 in the baroclinic model, especially in the vicinity of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}. Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia \citet{crean1988mathematical} and local adjustments were applied in both their barotropic and baroclinic models.  Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction although their domain did not include Puget Sound and the northern Strait of Georgia. \citet{foreman2004m} assimilated tide gauge harmonics to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.

>>>>

%overview oF SoG/Salish Sea models
The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion. %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers. Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome. They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004 in the baroclinic model, especially in the vicinity of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}. Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia \citep{crean1988mathematical}. Local adjustments to bottom friction were applied in both their barotropic and baroclinic models.  Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction although their domain did not include Puget Sound and the northern Strait of Georgia. \citet{foreman2004m} assimilated tide gauge harmonics to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.
